---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
zipfile = "activity.zip"
datafile = "activity.csv"

if( !file.exists(datafile) )
  unzip(zipfile)


rawdata <- read.csv(datafile)
rawdata$date <- as.POSIXlt(rawdata$date, format="%Y-%m-%d")

```

## What is mean total number of steps taken per day?

```{r}
library(plyr)
dailysum <- ddply(rawdata, .(date), summarize, total=sum(steps, na.rm=TRUE))
hist(dailysum$total, breaks=10,
  	main="Histogram of number of steps per day",
		xlab="Number of steps")

meandailysteps <- mean(dailysum$total, na.rm=TRUE)
mediandailysteps <- median(dailysum$total, na.rm=TRUE)


```

The mean number of steps per day is `r meandailysteps`.
The median number of steps per day is `r mediandailysteps`.


## What is the average daily activity pattern?

```{r}
intervalsum <- ddply(rawdata, .(interval), summarize, total=sum(steps, na.rm=TRUE))
plot(intervalsum$total, type="l")
maxstepinterval <- intervalsum$interval[which.max(intervalsum$total)]

```

The interval that contains the maximum number of steps is number `r maxstepinterval`.

## Imputing missing values

```{r}
numempty <- sum(is.na(rawdata$steps))

numdays = 61
intervalsumrep <- rep.int(t(intervalsum$total/numdays), numdays)
imputeddata <- rawdata
imputeddata$steps[is.na(rawdata$steps)] <- 
  intervalsumrep[is.na(rawdata$steps)]

dailysumimputed <- ddply(imputeddata, .(date), summarize, total=sum(steps, na.rm=TRUE))
hist(dailysumimputed$total, breaks=10,
		main="Histogram of number of steps per day",
		xlab="Number of steps")

meandailystepsimputed <- mean(dailysumimputed$total, na.rm=TRUE)
mediandailystepsimputed <- median(dailysumimputed$total, na.rm=TRUE)

```

The mean number of steps per day, including imputed values, is `r meandailystepsimputed`.
The median number of steps per day, including imputed values, is `r mediandailystepsimputed`.
The mean increases, but the median does not change.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
imputeddata$daytype <- as.factor(imputeddata$date$wday>0 & imputeddata$date$wday<6)


weekdaymean <- ddply(imputeddata[imputeddata$daytype==TRUE,], .(interval),
  			summarize, total=mean(steps, na.rm=TRUE))
weekendmean <- ddply(imputeddata[imputeddata$daytype==FALSE,], .(interval),
				summarize, total=mean(steps, na.rm=TRUE))

par(mfrow=c(2,1))
plot(weekdaymean$total, type="l", main="weekdays",
			 xlab="interval", ylab="number of steps")
plot(weekendmean$total, type="l", main="weekends",
			 xlab="interval", ylab="number of steps")

```